---
title: "Final Assessment: Return Transmission and Herding Behavior in the Semiconductor Sector"
subtitle: "Financial Econometrics – MSc in International Management, Finance"
author:
  - name: "Ludovic DELOT"
    email: "ludovic.delot‑bravo@rennes‑sb.com"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
format:
  html:
    theme: united            # Estilo visual cerulean, ajustable
    highlight: tango           # Resaltado de código
    toc: true                  # Tabla de contenido
    toc-location: left         # Ubicación de la tabla de contenido
    number-sections: false      # Numeración de secciones
    section-divs: true         # Divisiones por sección
    css: styles/custom.css     # Estilos adicionales a través de un CSS
    embed-resources: true      # Genera un archivo HTML autónomo
    html-math-method: katex    # Método para fórmulas matemáticas
lang: en
editor: visual                # Habilita editor visual en Quarto/RStudio
fontsize: 12pt                # Tamaño de fuente
geometry: margin=1in          # Márgenes en documento PDF
draft: false
---

## 

## II. DATA ACQUISITION & DESCRIPTION

### 2.1 Ticker Selection

| Ticker | Name                    | **Segme**nt   |
|--------|-------------------------|---------------|
| ASML   | ASML Holding            | Equipment     |
| AMAT   | Applied Materials       | Equipment     |
| LRCX   | Lam Research            | Equipment     |
| KLAC   | KLA Corp                | Equipment     |
| TER    | Teradyne                | Equipment     |
| ONTO   | Onto Innovation         | Equipment     |
| MKSI   | MKS Instruments         | Equipment     |
| ENTG   | Entegris                | Materials     |
| GLW    | Corning                 | Materials     |
| TSM    | Taiwan Semi (TSMC)      | Manufacturing |
| GFS    | GlobalFoundries         | Manufacturing |
| UMC    | United Microelectronics | Manufacturing |
| INTC   | Intel                   | Manufacturing |
| TXN    | Texas Instruments       | Manufacturing |
| MU     | Micron                  | Manufacturing |
| ON     | ON Semi                 | Manufacturing |
| MCHP   | Microchip Tech          | Manufacturing |
| STM    | STMicroelectronics      | Manufacturing |
| NVDA   | NVIDIA                  | Design        |
| AMD    | AMD                     | Design        |
| AVGO   | Broadcom                | Design        |
| QCOM   | Qualcomm                | Design        |
| MRVL   | Marvell Tech            | Design        |
| ADI    | Analog Devices          | Design        |
| NXPI   | NXP Semiconductors      | Design        |
| ASX    | ASE Technology          | OSAT          |
| AMKR   | Amkor Technology        | OSAT          |
| ARW    | Arrow Electronics       | Distributor   |
| AVT    | Avnet                   | Distributor   |
| SOXX   | iShares Semi ETF        | Benchmark     |
| SMH    | VanEck Semi ETF         | Benchmark     |
| XSD    | SPDR S&P Semi ETF       | Benchmark     |
| \^GSPC | S&P 500 Index           | Market        |
| \^IXIC | NASDAQ Composite        | Market        |
| QQQ    | Invesco QQQ             | Market        |
| \^SOX  | PHLX Semiconductor      | Market        |
| \^VIX  | CBOE VIX                | Risk          |
| \^TNX  | US 10Y Yield            | Rates         |
| \^IRX  | 13W T-Bill              | Rates         |
| UUP    | US Dollar Bullish (UUP) | Rates         |
| CL=F   | WTI Crude Oil (Front)   | Costs         |
| HG=F   | COMEX Copper (Front)    | Costs         |
| AAPL   | Apple                   | Downstream    |
| TSLA   | Tesla                   | Downstream    |

```{r}
#| label: setup-packages
#| include: false
#| message: false
#| warning: false

if (!requireNamespace("pacman", quietly = TRUE)) {
install.packages("pacman")
}

pacman::p_load(
tidyverse, tidyquant, lubridate, janitor,
moments, tseries, knitr, ggplot2,
rugarch, FinTS, lmtest, sandwich,
car, PerformanceAnalytics
)

```

```{r}
#| include: false
# we create a table with tickers and their corresponding names and segments (like 'Equipment', 'Design', etc.)
tickers_tbl <- tibble::tibble(
  ticker  = c(
    # Equipment / Tools
    "ASML","AMAT","LRCX","KLAC","TER","ONTO","MKSI",
    # Materials / Inputs
    "ENTG","GLW",
    # Manufacturing (Foundry/IDM/Memory/Power)
    "TSM","GFS","UMC","INTC","TXN","MU","ON","MCHP","STM",
    # Design / Fabless
    "NVDA","AMD","AVGO","QCOM","MRVL","ADI","NXPI",
    # OSAT / Packaging
    "ASX","AMKR",
    # Distributors
    "ARW","AVT",
    # Benchmarks (ETFs)
    "SOXX","SMH","XSD",
    # --- Controls we add ---
    # Market & Tech
    "^GSPC","^IXIC","QQQ","^SOX",
    # Risk
    "^VIX",
    # Rates & Dollar
    "^TNX","^IRX","UUP",
    # Costs
    "CL=F","HG=F",
    # Downstream demand proxies
    "AAPL","TSLA"
  ),
  name    = c(
    # Equipment / Tools
    "ASML Holding","Applied Materials","Lam Research","KLA Corp","Teradyne","Onto Innovation","MKS Instruments",
    # Materials / Inputs
    "Entegris","Corning",
    # Manufacturing
    "Taiwan Semi (TSMC)","GlobalFoundries","United Microelectronics","Intel","Texas Instruments","Micron","ON Semi","Microchip Tech","STMicroelectronics",
    # Design / Fabless
    "NVIDIA","AMD","Broadcom","Qualcomm","Marvell Tech","Analog Devices","NXP Semiconductors",
    # OSAT
    "ASE Technology","Amkor Technology",
    # Distributors
    "Arrow Electronics","Avnet",
    # Benchmarks
    "iShares Semi ETF","VanEck Semi ETF","SPDR S&P Semi ETF",
    # --- Controls we add ---
    # Market & Tech
    "S&P 500 Index","NASDAQ Composite","Invesco QQQ","PHLX Semiconductor",
    # Risk
    "CBOE VIX",
    # Rates & Dollar
    "US 10Y Yield","13W T-Bill","US Dollar Bullish (UUP)",
    # Costs
    "WTI Crude Oil (Front)","COMEX Copper (Front)",
    # Downstream
    "Apple","Tesla"
  ),
  segment = c(
    # Equipment / Tools
    rep("Equipment", 7),
    # Materials / Inputs
    rep("Materials", 2),
    # Manufacturing
    rep("Manufacturing", 9),
    # Design / Fabless
    rep("Design", 7),
    # OSAT
    rep("OSAT", 2),
    # Distributors
    rep("Distributor", 2),
    # Benchmarks
    rep("Benchmark", 3),
    # --- Controls we add ---
    # Market & Tech
    rep("Market", 4),
    # Risk
    "Risk",
    # Rates & Dollar
    rep("Rates", 3),
    # Costs
    rep("Costs", 2),
    # Downstream
    rep("Downstream", 2)
  )
)
```

### 2.2 Setting Data Range

```{r}
#| echo: false
#| warning: false
#we set date range for the data
from_date <- as.Date("2018-01-01") #start
to_date   <- as.Date("2025-10-24")   #end
```

### 2.3 Downloading the Data from YF

```{r}
#| echo: false
#| warning: false
#we download daily stock price data from Yahoo Finance for all tickers #and this function gets the stock data for the tickers we defined earlier from Yahoo
prices_raw <- tidyquant::tq_get(
  x    = tickers_tbl$ticker,   # the list of tickers
  get  = "stock.prices",       # we want daily stock prices
  from = from_date,            # from the start date
  to   = to_date               # until the end date
)
```

```{r}
#| echo: false
#| warning: false
#we combine the stock prices with the ticker names and segments
df_1_raw <- prices_raw %>%
  dplyr::rename(ticker = symbol) %>%               #rename 'symbol' to 'ticker'
  dplyr::left_join(tickers_tbl, by = "ticker") %>% #add names and segments
  dplyr::mutate(source = "yahoo") %>%              #record the source as Yahoo
  dplyr::select(                                   #select only useful columns
    ticker, name, segment, date, open, high, low, close, adjusted, volume, source
  ) %>%
  dplyr::arrange(ticker, date)                     #arrange data by ticker and date
```

### 2.5 Exploratory Data Analysis

```{r}
#| echo: false
#| warning: false
#first we use glimpse to get a basic summary
glimpse(df_1_raw)            

#count how many unique tickers there are
cat("Total number of unique tickers: ", dplyr::n_distinct(df_1_raw$ticker), "\n")

#then we create the log returns

df_2_returns <- df_1_raw %>%
  dplyr::group_by(ticker) %>%              #group data by ticker
  dplyr::arrange(date, .by_group = TRUE) %>%  #sort by date within each ticker group
  dplyr::mutate(ret_log = log(adjusted / dplyr::lag(adjusted))) %>% 
  #calculate log returns
  dplyr::ungroup()                            #ungroup after the calculation

df_2_returns <- df_2_returns %>%
  dplyr::filter(!is.na(ret_log))

#eda

#summary
summary_tbl <- df_2_returns %>%
  dplyr::group_by(ticker, name, segment) %>%  #group by ticker, name, and segment
  dplyr::summarise(
    n_obs   = sum(!is.na(ret_log)),         #number of observations but with not na
    mean    = mean(ret_log, na.rm = TRUE),  #mean return
    median  = median(ret_log, na.rm = TRUE), #median return
    sd      = sd(ret_log, na.rm = TRUE),    #standard deviation (volatility)
    min     = min(ret_log, na.rm = TRUE),   #minimum return
    max     = max(ret_log, na.rm = TRUE),   #maximum return
    skew    = moments::skewness(ret_log, na.rm = TRUE),  #skewness (measuring asymmetry)
    kurt    = moments::kurtosis(ret_log, na.rm = TRUE),  #kurtosis (measuring extreme values)
    jb_p    = tryCatch(tseries::jarque.bera.test(stats::na.omit(ret_log))$p.value,
                       error = function(e) NA_real_),   #jarquebera test for normality
    .groups = "drop"  #drop the grouping after summarizing
  ) %>%
  dplyr::arrange(segment, ticker)       #arrange by segment and ticker

#show the descriptive statistics table
knitr::kable(summary_tbl, caption = "Descriptive statistics by asset (log returns)")

#top 10 by volatility (highest standard deviation)
print("Top 10 tickers by volatility (highest sd")
top_volatility <- summary_tbl %>% dplyr::arrange(dplyr::desc(sd)) %>% head(10)
knitr::kable(top_volatility)

#op 10 by negative skew (lowest skew)
print("Top 10 tickers by negative skew (lowest skew)")
top_neg_skew <- summary_tbl %>% dplyr::arrange(skew) %>% head(10)
knitr::kable(top_neg_skew)

#ttop 10 by highest kurtosis (fat tails)
print("Top 10 tickers by highest kurtosis")
top_kurtosis <- summary_tbl %>% dplyr::arrange(dplyr::desc(kurt)) %>% head(10)
knitr::kable(top_kurtosis)

returns_wide <- df_2_returns %>%
dplyr::select(date, ticker, ret_log) %>% # Select the relevant columns
tidyr::spread(key = ticker, value = ret_log) %>% # Convert to wide format
tidyr::drop_na()
```

## III. PROBABILITY & DISTRIBUTION ANALYSIS

### 3.1 Select Variable (NVIDIA log returns)

```{r}
#| echo: false
#| message: false
#| warning: false
nvda_ret <- df_2_returns %>%
dplyr::filter(ticker == "NVDA", !is.na(ret_log)) %>%
dplyr::select(date, ret_log) %>%
dplyr::mutate(year = lubridate::year(date))

#quick summary of observations for nvda

cat("✅ Observations for NVDA:", nrow(nvda_ret), "\n")
print(summary(nvda_ret$ret_log))  # summary of log returns

```

### 3.2 Empirical PDF and CDF

```{r}
#| echo: false
#| message: false
#| warning: false

# empirical probability density function (pdf)

nvda_density <- stats::density(nvda_ret$ret_log, na.rm = TRUE)
dens_df <- tibble::tibble(x = nvda_density$x, y = nvda_density$y)

# plot histogram with overlaid density curve

ggplot2::ggplot(nvda_ret, ggplot2::aes(x = ret_log)) +
ggplot2::geom_histogram(ggplot2::aes(y = ..density..), bins = 60, fill = "#1DB954", alpha = 0.5, color = "white") +
ggplot2::geom_line(data = dens_df, ggplot2::aes(x, y), color = "darkgreen", linewidth = 0.9) +
ggplot2::labs(title = "nvidia — empirical pdf (log returns)", x = "log return", y = "density") +
ggplot2::theme_minimal()

# empirical cumulative distribution function (cdf)

ggplot2::ggplot(nvda_ret, ggplot2::aes(x = ret_log)) +
ggplot2::stat_ecdf(geom = "step", color = "#1DB954", linewidth = 0.9) +
ggplot2::labs(title = "nvidia — empirical cdf (log returns)", x = "log return", y = "cumulative probability") +
ggplot2::theme_minimal()

```

### 3.3 Normality Tests and QQ plots

```{r}
#| echo: false
#| message: false
#| warning: false
#performing normality tests
n_sub     <- min(5000, nrow(nvda_ret))
shapiro_p <- stats::shapiro.test(nvda_ret$ret_log[1:n_sub])$p.value
jb_p      <- tseries::jarque.bera.test(stats::na.omit(nvda_ret$ret_log))$p.value
skew_val  <- moments::skewness(nvda_ret$ret_log)
kurt_val  <- moments::kurtosis(nvda_ret$ret_log)

#print normality test results

cat("\n### normality tests for nvidia log returns ###\n")
cat("shapiro p-value:", formatC(shapiro_p, format = "e", digits = 2), "\n")
cat("jarque-bera p-value :", formatC(jb_p, format = "e", digits = 2), "\n")
cat("skewness:", round(skew_val, 3), " | kurtosis:", round(kurt_val, 3), "\n")

#qq plot to visually inspect normality

stats::qqnorm(nvda_ret$ret_log, main = "qq plot for nvda log returns vs normal")
stats::qqline(nvda_ret$ret_log, col = "blue", lwd = 2)

```

### 3.4 Probability Analysis

```{r}
#| echo: false
#| message: false
#| warning: false
thresholds_log <- c(0, log(1.01), log(1.02), log(1.03), log(1.04))
names(thresholds_log) <- c("positive", "1%", "2%", "3%", "4%")

#calculate empirical probabilities for each threshold

probs_log <- sapply(thresholds_log, function(th)
mean(nvda_ret$ret_log > th, na.rm = TRUE)
)

#display probabilities as a table

prob_tbl_log <- tibble::tibble(
threshold_label = names(thresholds_log),
simple_return   = c("0%", "1%", "2%", "3%", "4%"),
log_threshold   = round(thresholds_log, 5),
probability     = round(probs_log * 100, 2)
)

cat("\n### empirical probabilities (log returns) — p(x > threshold) ###\n")
print(prob_tbl_log)

```

```{r}
#| echo: false
#| message: false
#| warning: false
# create histogram and density plot
ggplot(nvda_ret, aes(x = ret_log)) +
  geom_histogram(aes(y = ..density..), bins = 60, fill = "#1DB954", alpha = 0.5, color = "white") +  # Histogram
  geom_density(color = "darkgreen", linewidth = 0.9) +  # Density line
  labs(title = "NVIDIA — Empirical PDF (Log Returns)", x = "Log Return", y = "Density") +
  
  #add vertical lines for specific thresholds (1%, 2%, 3%)
  geom_vline(data = prob_tbl_log[prob_tbl_log$simple_return %in% c("1%", "2%", "3%"),], 
             aes(xintercept = log_threshold, color = simple_return), 
             linetype = "dashed", size = 1) +
  
  #define colors for the lines in the legend
  scale_color_manual(values = c("1%" = "blue", "2%" = "red", "3%" = "green")) +
  
  #position the legend at the top
  theme(legend.position = "top", legend.title = element_blank())


```

## IV. REGRESSION ANALYSIS

```{r}
#| echo: false
#| message: false
#| warning: false
#| 

#rename etf without ^
if ("^SOX" %in% names(returns_wide)) {
  returns_wide <- returns_wide %>% rename(SOX = `^SOX`)
}

#relevant tickers selection
tickers <- c(
  "NVDA","ADI","AMAT","AMD","ASML","AVGO","ENTG","INTC","KLAC","LRCX",
  "MCHP","MRVL","MU","NXPI","ON","ONTO","QCOM","SMH","SOXX","TER",
  "TSLA","TSM","TXN","XSD","SOX"
)

#ensure only selected tickers are in returns_wide
tickers <- intersect(tickers, names(returns_wide))

#automatically calculate log-returns
returns_log <- returns_wide %>%
  arrange(date) %>%
  mutate(across(all_of(tickers),
                ~ log(.x / lag(.x)),
                .names = "{.col}_logreturn")) %>%
  drop_na()

#dataset with only _logreturn columns
data_log <- returns_log %>%
  select(ends_with("_logreturn")) %>%
  drop_na()

#ensure NVDA_logreturn is present
stopifnot("NVDA_logreturn" %in% names(data_log))
cat("✅ Log-returns generated for", ncol(data_log), "tickers.\n")

#PCAS
y <- data_log$NVDA_logreturn
X <- data_log %>% select(-NVDA_logreturn)

#scale the features for PCA
X_scaled <- scale(X)

#perform PCA
pca_res  <- prcomp(X_scaled, center = TRUE, scale. = TRUE)

#extract the first 6 principal components
pcs6 <- as.data.frame(pca_res$x[, 1:6, drop = FALSE])
names(pcs6) <- paste0("PC", 1:6)

#show explained variance
var_exp <- (pca_res$sdev^2) / sum(pca_res$sdev^2)
cat("\nExplained variance (6 PCs):", round(sum(var_exp[1:6])*100, 2), "%\n")

#OLS with Robust Errors
df_ols <- bind_cols(NVDA_logreturn = y, pcs6)
ols_fit <- lm(NVDA_logreturn ~ ., data = df_ols)

cat("\n--- OLS (NVDA_logreturn ~ 6 PCs) ---\n")
print(summary(ols_fit))

#robust Standard Errors HC3
hc3_se <- vcovHC(ols_fit, type = "HC3")
print(coeftest(ols_fit, vcov = hc3_se))

#breusch-Pagan Test for Heteroscedasticity
cat("\n--- Breusch-Pagan Test for Heteroscedasticity ---\n")
print(bptest(ols_fit))

#normality of Residuals ---
cat("\n--- Normality of Residuals ---\n")
resid_ols <- residuals(ols_fit)
print(shapiro.test(sample(resid_ols, min(5000, length(resid_ols)))))
print(jarque.bera.test(resid_ols))

#LOGIT (Yes/No Return)
#create binary variable: 1 if return > 0, else 0
NVDA_up <- as.integer(y > 0)
df_logit <- bind_cols(NVDA_up = NVDA_up, pcs6)

logit_fit <- glm(NVDA_up ~ ., data = df_logit, family = binomial())

cat("\n--- LOGIT (Positive Return ~ 6 PCs) ---\n")
print(summary(logit_fit))

#make predictions
p_pred <- predict(logit_fit, type = "response")
pred <- ifelse(p_pred >= 0.5, 1, 0)

#confusion matrix and accuracy
tab <- table(Truth = df_logit$NVDA_up, Pred = pred)
acc <- mean(pred == df_logit$NVDA_up)
cat("\nConfusion Matrix (Yes/No Return):\n")
print(tab)
cat("Accuracy (cutoff=0.5):", round(acc, 3), "\n")

cat("\nProportion of positive returns in the dataset:", round(mean(df_logit$NVDA_up), 3), "\n")
cat("\n✅ Model completed: OLS + LOGIT on 6 PCs (auto-generated log-returns).\n")



```

```{r}
#extract residuals and fitted values from the OLS model
resid_ols   <- residuals(ols_fit)
fitted_ols  <- fitted(ols_fit)

#for error fixing
time_index <- seq_along(resid_ols)

#fitted value vs residualts
ggplot(
  data.frame(Fitted = fitted_ols, Residuals = resid_ols),
  aes(x = Fitted, y = Residuals)
) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(
    title = "Fitted Values vs Residuals",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()
 
#residuals over time
ggplot(
  data.frame(Observation = time_index, Residuals = resid_ols),
  aes(x = Observation, y = Residuals)
) +
  geom_line(color = "steelblue") +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(
    title = "Residuals Over Time",
    x = "Observation",
    y = "Residuals"
  ) +
  theme_minimal()

#scale location
ggplot(
  data.frame(Fitted = fitted_ols,
             SqrtAbsResid = sqrt(abs(resid_ols))),
  aes(x = Fitted, y = SqrtAbsResid)
) +
  geom_point(alpha = 0.6) +
  geom_smooth(method = "loess", se = FALSE, color = "red") +
  labs(
    title = "Scale–Location Plot",
    x = "Fitted Values",
    y = "√|Residuals|"
  ) +
  theme_minimal()

#qq plot
ggplot(
  data.frame(Residuals = resid_ols),
  aes(sample = Residuals)
) +
  geom_qq() +
  geom_qq_line() +
  labs(
    title = "Q–Q Plot of Residuals",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_minimal()

#normality test
cat("\n--- Shapiro–Wilk Normality Test ---\n")
print(shapiro.test(resid_ols))

cat("\n--- Jarque–Bera Normality Test ---\n")
print(jarque.bera.test(resid_ols))




```

```{r}


#residuals and fitted values
resid_ols  <- residuals(ols_fit)
fitted_ols <- fitted(ols_fit)

#residuals vs Fitted
ggplot(
  data.frame(Fitted = fitted_ols, Residuals = resid_ols),
  aes(x = Fitted, y = Residuals)
) +
  geom_point(alpha = 0.6) +
  geom_hline(yintercept = 0, color = "red", linewidth = 1) +
  labs(
    title = "Residuals vs Fitted",
    x = "Fitted values",
    y = "Residuals"
  ) +
  theme_minimal()

#cook's Distance plot
cook_vals <- cooks.distance(ols_fit)

ggplot(
  data.frame(Observation = seq_along(cook_vals), CookD = cook_vals),
  aes(x = Observation, y = CookD)
) +
  geom_bar(stat = "identity", alpha = 0.7) +
  geom_hline(yintercept = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Cook's Distance",
    x = "Observation",
    y = "Cook's D"
  ) +
  theme_minimal()

#VIF bar plot
vif_vals <- vif(ols_fit)
vif_df <- data.frame(
  Variable = names(vif_vals),
  VIF      = as.numeric(vif_vals)
)

ggplot(vif_df, aes(x = reorder(Variable, VIF), y = VIF)) +
  geom_col(alpha = 0.7) +
  coord_flip() +
  labs(
    title = "Variance Inflation Factors (VIF)",
    x = "Predictor",
    y = "VIF"
  ) +
  theme_minimal()

```

## V. HERDING

```{r}
#filter log returns for the tickers of interest (NVDA, QCOM, TSM, AMD)

#we use df_2_returns to filter the data for the selected tickers and calculate the log returns.

tickers <- c("NVDA", "QCOM", "TSM", "AMD")

#we filter the data to select only the required tickers and remove any rows with missing values

log_returns <- df_2_returns %>%
dplyr::filter(ticker %in% tickers) %>%  #filter for selected tickers
dplyr::select(date, ticker, ret_log) %>%  #select date, ticker, and log returns
tidyr::spread(key = ticker, value = ret_log) %>%  #spread the data to a wide format (tickers as columns)
tidyr::drop_na()  #drop any rows with missing values (after spreading)

#CSAD measures the dispersion of individual stock returns around the average market return.
#define the CSAD calculation function

calculate_CSID <- function(return_data) {

#calculate the average market return (Rm) by taking the row means of returns

Rm <- rowMeans(return_data, na.rm = TRUE)

#calculate the absolute deviation of each stock's return from the market return

CSAD <- rowMeans(abs(return_data - Rm), na.rm = TRUE)
return(CSAD)
}

#apply the CSAD function to the log returns data (excluding the date column)

CSAD_values <- calculate_CSID(log_returns[,-1])  #exclude the date column
#create a new dataframe for the CSAD analysis
CSAD_df <- data.frame(Date = log_returns$date, CSAD = CSAD_values, Rm = rowMeans(log_returns[,-1], na.rm = TRUE))
CSAD_df$Rm2 <- CSAD_df$Rm^2  #we addd the quadratic term (Rm^2) to capture non-linearity
#nw, we fit a linear model that includes both the market return (Rm) and the quadratic term (Rm^2)
model <- lm(CSAD ~ Rm + Rm2, data = CSAD_df)
#display the summary of the model to check the coefficients and significance
summary(model)
#visualize the Results
#plot CSAD against the market return (Rm) with the fitted regression line
plot(CSAD_df$Rm, CSAD_df$CSAD, main = "CSAD vs. Rm", xlab = "Market Return (Rm)", ylab = "CSAD", col = "blue")
abline(model, col = "red")  # Add the regression line


```

## VI. TIME SERIES MODELING

```{r}
soxx_ret <- df_2_returns %>%
dplyr::filter(ticker == "SOXX") %>%
dplyr::select(date, ret_log) %>%
tidyr::drop_na()
#check for volatility clustering
#we use the bbox-ljung test and ARCH test to check for autocorrelation and volatility clustering in returns
Box.test(soxx_ret$ret_log, lag = 10, type = "Ljung-Box")
ArchTest(soxx_ret$ret_log, lags = 10)
#define and fit EGARCH(1,1) model
spec_egarch <- ugarchspec(
variance.model = list(model = "eGARCH", garchOrder = c(1, 1)),
mean.model     = list(armaOrder = c(1, 1), include.mean = TRUE),
distribution.model = "std"
)
#fit the model to the SOXX returns
fit_egarch <- ugarchfit(spec = spec_egarch, data = soxx_ret$ret_log)
#show model summary
show(fit_egarch)
#visualize key plots from the EGARCH model
par(mfrow = c(2, 2), mar = c(2, 2, 2, 2))
plot(fit_egarch, which = 1)  #Conditional variance plot
plot(fit_egarch, which = 2)  #Standardized residuals plot
plot(fit_egarch, which = 3)  #ACF of residuals plot
plot(fit_egarch, which = 9)  #QQ plot of residuals

```

## VII. FACTOR ANALYSIS / PCA

```{r}
###alredy in the regression part
```

## 
